Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2023-02-24 analysis/paper/paper.qmd
add new citation & create png file for map

Tip cross-sectional area (TCSA) of stone artifacts has been used as a ballistically relevant standard to probabilistically discriminate different likely use classes of projectile armatures, such as spear-thrower (a.k.a. atlatl) dart tips, un-poisoned arrow tips and large stabbing/thrusting spears [@sisk2011african; @lombard2021variation; @lombard2020tip; @lombard2021did; @o2018origins; @lombard2022standardized]. It is critical to note that the TCSA metric alone cannot unambiguously determine artifact function; it only suggests a best-fit ballistic probability for the points if they were hafted as weapon tips. The TCSA metric represents the part of the tool that cuts the target's hide and is related to weapon flight and penetration dynamics [@hughes1998getting; @sitton2020tip; @lombard2021variation]. This method was first proposed by @hughes1998getting and validated by @shea2006origins through the comparison of archaeological examples with ethnographically collected samples of known use. @shea2009impact also applied this approach to compare projectile points from Africa, the Levant, and Europe, claiming that projectile weapons first appeared in Africa.

@@

>>>>

Tip cross-sectional area (TCSA) of stone artifacts has been used as a ballistically relevant standard to probabilistically discriminate different likely use classes of projectile armatures, such as spear-thrower (a.k.a. atlatl) dart tips, un-poisoned arrow tips and large stabbing/thrusting spears [@sisk2011african; @lombard2021variation; @lombard2020tip; @lombard2021did; @o2018origins; @lombard2022standardized; @metz2023bow]. It is critical to note that the TCSA metric alone cannot unambiguously determine artifact function; it only suggests a best-fit ballistic probability for the points if they were hafted as weapon tips. The TCSA metric represents the part of the tool that cuts the target's hide and is related to weapon flight and penetration dynamics [@hughes1998getting; @sitton2020tip; @lombard2021variation]. This method was first proposed by @hughes1998getting and validated by @shea2006origins through the comparison of archaeological examples with ethnographically collected samples of known use. @shea2009impact also applied this approach to compare projectile points from Africa, the Levant, and Europe, claiming that projectile weapons first appeared in Africa.

@@

# ggsave(here::here("analysis/figures/fig-map.png"),
#        width = 4,
#        height = 4.5,
#        units = "in")
```
